### 2013-11-04 proposal/proposal.tex
initial commit: added README, proposal in progress

\documentclass{proc}
\usepackage[colorlinks=true]{hyperref}

\begin{document}

\title{Generating Street Maps with Tensor Fields}
\author{William Clarkson, Marcella Hastings, Nathaniel Tenczar}

\maketitle

\section{Problem Statement}
Our goal is to generate realistic street maps which based on local geography
such as water boundaries, and which can be easily shaped by user-defined
control points to quickly and easily achieve a desired layout. This has been
achieved in \cite{wonka}. However, their implementation is roughly 100,000
lines of C++, which the authors found difficult to understand due to a lack
of modularity. We propose that by implementing the core functionality of this
pipeline in a clear and modular way, we will enable future contributors
to comprehend and build more easily upon this work.

\section{Context}
We will implement this pipeline in Haskell and make full use of the functional
programming paradigm to accomplish the goals stated above. As many of the
components of the pipeline are defined mathematically, a functional
language like Haskell will lend itself well to expressing them.

\section{Overview of Solution}
As described in \cite{wonka}, we will generate a tensor field based on a set
of constraints (including water boundaries and user-defined control points
indicating street orientation). We will calculate the major and minor
eigenvectors of this tensor field, and trace hyperstreamlines based on
the eigenvectors to determine the locations of streets. We will then calculate
intersections between these hyperstreamlines to build a graph of streets. We
will then generate a visual representation of this street layout.

\section{Obstacles}
At present, we are unfamiliar with tensor fields and their associated
operations. We identify this as the primary challenge to the completion of this
project. However, we are confident that, with our background knowledge of
mathematics and the wealth of available resources on the topic, we will be able
to implement the necessary representations and operations of tensor fields to
solve the problem as a whole.

\section{Definition of Success}
Our project will be a success if our implementation of the work of Chen et al.
allows a user to provide simple constraints from which an image of a realistic
street layout is generated which is influenced by the provided constraints. If
time permits, we would like to provide an interactive interface in which users
can intuitively add and edit constraints in real time.

\begin{thebibliography}{9}
\bibitem{wonka}
    Chen, G., Esch, G., Wonka, P., Müller, P., Zhang, E.. 2008. \href{http://www.google.com/url?q=http%3A%2F%2Fpeterwonka.net%2FPublications%2Fpdfs%2F2008.SG.Chen.InteractiveProceduralStreetModeling.pdf&sa=D&sntz=1&usg=AFQjCNGp-DPZdfTyZLHmKVeefan3jltzzw}{Interactive Procedural Street Modeling}. ACM Transactions on Graphics.
\bibitem{tensorviz}
    Komura, Taku.  \href{http://www.google.com/url?q=http%3A%2F%2Fwww.inf.ed.ac.uk%2Fteaching%2Fcourses%2Fvis%2Flecture_notes%2Flecture14.pdf&sa=D&sntz=1&usg=AFQjCNGytegkIGcZJd_JM-3pzlilzPVH7g}{Tensor Visualization Lecture Notes}. Institute for Perception, Action \& Behaviour, School of Informatics, University of Edinburgh.
\bibitem{tensors}
    Delmarcelle, T.; Hesselink, L. 2006. \href{http://www.google.com/url?q=http%3A%2F%2Fwww.inf.ethz.ch%2Fpersonal%2Fpeikert%2FSciVis%2FLiterature%2FDelmarcelleHesselink94.pdf&sa=D&sntz=1&usg=AFQjCNGIb52UifdoK_NB0oB8sU_r_rebBQ}{The Topology of Symmetric, Second-Order Tensor Fields}. Visualization and Processing of Tensor Fields.

\end{thebibliography}

\end{document}

### 2013-11-04 proposal/proposal.tex
changed 'the authors' to 'we'
Adding where constraints come from

@@
 Our goal is to generate realistic street maps which based on local geography
-such as water boundaries, and which can be easily shaped by user-defined
+such as water boundaries, and which can be shaped by user-defined
 control points to quickly and easily achieve a desired layout. This has been
 achieved in \cite{wonka}. However, their implementation is roughly 100,000
-lines of C++, which the authors found difficult to understand due to a lack
+lines of C++, which we found difficult to understand due to a lack
 of modularity. We propose that by implementing the core functionality of this
@@
 solve the problem as a whole.
+\par
+We do not anticipate developing a user interface as our first priority, so users
+will specify constraints by writing a hardcoded file with their desired 
+specifications.
 

### 2013-12-10 paper/paper.tex
added json test files and paper template

\documentclass[twocolumn]{article}
% \usepackage{bibtex}

\title{Modularizing Street Map Generation}
\author{William Clarkson \and Marcella Hastings \and Nathaniel Tenczar}
\date{\today}

\begin{document}
\maketitle

\begin{abstract}
    Abstract will go here. \cite{chen}
\end{abstract}

\section{Introduction}

\section{Existing Work}

\section{Our Goals}

\section{Mathematical Background}

\section{Pipeline}

\section{Implementation Details}

\section{Assessment of our Work}

\section{Contributions}

\bibliographystyle{plain}
\bibliography{references}

\end{document}

### 2013-12-11 Streamline.hs
fixed streamline tracing to actually pass through nearest neighbor and do major-major and minor-minor streamline intersections

@@ -1,6 +1,6 @@
 module Streamline (PlacementMethod, traceLines) where
 
-import TensorField
+import TensorField as TF
 import Vector2 (Vector2 (Vector2))
 import SVGWriter
 import NearestNeighbor (Storage)
@@ -39,10 +39,10 @@ sampleList xs =
                         []     -> []
   in everyNth 10 xs
 
-placeSeeds :: (VectorField, VectorField) -> [Streamline] -> Int ->
-  Float -> Float -> [Streamline]
-placeSeeds _         sls 0 _  _  = sls
-placeSeeds (maj,min) sls n fw fh =
+placeSeeds :: (Num a) => (VectorField, VectorField) -> Storage a -> [Streamline] -> Int ->
+  Float -> Float -> ([Streamline], Storage a)
+placeSeeds _         nn sls 0 _  _  = (sls, nn)
+placeSeeds (maj,min) nn sls n fw fh =
   let existPts = concatMap sampleList sls
       trialPts = randomSeeds 10 fw fh
       dist (Vector2 x y) (Vector2 x2 y2) = (x-x2)*(x-x2) + (y-y2)*(y-y2)
@@ -50,24 +50,37 @@ placeSeeds (maj,min) sls n fw fh =
                       [] -> 2*fw  -- this is evil
                       _  -> minimum (map (dist v) existPts)
       seed     = Util.argmax closestPt trialPts
-      nn       = NN.new fw fh 4
       majSl    = streamlineFromEvs maj nn seed fw fh
       minSl    = streamlineFromEvs min nn seed fw fh
-  in placeSeeds (maj,min) (majSl:minSl:sls) (n - 1) fw fh
+      nn'      = TF.addStreamlineToNN majSl nn
+      nn''     = TF.addStreamlineToNN minSl nn'
+  in placeSeeds (maj,min) nn'' (majSl:minSl:sls) (n - 1) fw fh
 
-placeSeedsImproved ::
-  (VectorField, VectorField) -> [Streamline] -> Int -> Float -> Float -> [Streamline]
-placeSeedsImproved _         sls 0 _  _  = sls
-placeSeedsImproved (maj,min) sls n fw fh =
+placeSeedsImproved :: (Num a) =>
+  (VectorField, VectorField) -> Storage a -> [Streamline] -> Int -> Float -> Float -> ([Streamline], Storage a)
[…]

### 2013-12-12 Tensor.hs
simplified eigenvalue calculation based on the fact that tensor matrix is traceless

@@ -44,8 +44,10 @@ isReal (Tensor a b c) =
 -- returns a tuple of the (larger, smaller) eigenvectors of a tensor
 eigenvalues :: Tensor -> (Float, Float)
 eigenvalues (Tensor a b d) = 
-  let evals f = (f (a + d) (sqrt (4*b*b + (a - d)*(a - d)))) / 2
-  in  (evals (+), evals (-))
+--  let evals f = (f (a + d) (sqrt (4*b*b + (a - d)*(a - d)))) / 2
+  let eval = sqrt (a*a + b*b) 
+  in  (eval, (-1)*eval)
+-- in  (evals (+), evals (-))
 
 -- returns a tuple of the (major, minor) eigenvectors for a tensor
 eigenvectors :: Tensor -> (V2.Vector2, V2.Vector2)

### 2013-12-12 paper/paper.tex
fixed streamline tracing to actually pass through nearest neighbor and do major-major and minor-minor streamline intersections
added a bunch of content from the google doc to the paper and typeset most equations

@@
 \documentclass[twocolumn]{article}
-% \usepackage{bibtex}
+\usepackage{amsmath}
 
-\title{Modularizing Street Map Generation}
+\title{Modularizing Street Map Generation \\
+    \vspace{8pt} \large An Adventure in Tensors}
 \author{William Clarkson \and Marcella Hastings \and Nathaniel Tenczar}
 \date{\today}
+
+\newcommand{\sqmat}[4]{\ensuremath{
+    \left(\begin{array}{cc}
+        #1 & #2 \\
+        #3 & #4
+    \end{array}\right)}}
+\newcommand{\mkvec}[2]{\ensuremath{
+    \left(\begin{array}{c}
+        #1 \\
+        #2 \\
+    \end{array}\right)}}
+\newcommand{\pt}{\textbf{p}}
+\newcommand{\todo}[1]{\begin{center}\fbox{\parbox{150pt}{#1}}\end{center}}
 
@@
 \begin{abstract}
-    Abstract will go here. \cite{chen}
 \end{abstract}
@@
 \section{Mathematical Background}
+A tensor $t$ is a *magical thing* represented by a 2x2 symmetric, traceless
+matrix:
+\[
+    \sqmat{\cos{2\theta}}{\sin{2\theta}}{\sin{2\theta}}{-\cos{2\theta}}
+    = \sqmat{a}{b}{b}{-a}
+\]
+where $R$ is the magnitude and $\theta$ is the direction, and
+$R\geq0 \textrm{ and } \theta\in[0,2\pi)$. Since the matrix is symmetric and
+traceless, it is only necessary to store two values, $a$ and $b$ to encode
+the tensor.
+
+The major eigenvectors of $t$ are
+\[
+    \left\{
+        \lambda\mkvec{\cos{\theta}}{\sin{\theta}} ~|~ \lambda \neq 0
+    \right\}
+\]
+The minor eigenvectors are
+\[
+    \left\{
+        \lambda\mkvec
+                {\cos{(\theta+\frac{\pi}{2})}}
+                {\sin{(\theta+\frac{\pi}{2})}}
+        ~|~ \lambda \neq 0
+    \right\}
+\]
+A tensor field $T$ is a function that maps from 2D points to tensors. In the
+context of this project, a tensor is simply a tool we use to determine the
+direction of a field at a point. This direction is found by calculating the
+eigenvectors of the tensors. Major eigenvectors are parallel to the field;
+minor eigenvectors are perpendicular.
+
+Another important concept is the hyperstreamline. This is an effective way to
+visualize tensor fields. It is created by drawing a line through an eigenvector
+field such that the line is tangent to every eigenvector along its path. In
+this setting, there are major and minor hyperstreamlines that correspond to the
+major and minor eigenvectors. The major and minor eigenvectors of a tensor are
+guaranteed to be perpendicular for any tensor, unless it is a degenerate point
+in the tensor field ($T(\bf p)=0$), for which eigenvectors cannot be
+calculated. This property leads to the generation of perpendicular street
+intersections, a common feature of real-world road networks, especially in
+cities.
 
@@
 
+\subsection{User-Generated Constraints}
+
+\subsection{Tensor Field Generation}
+From each constraint, a tensor field is defined across the entire domain; this
+is called a basis field. The basis fields are then combined to produce a tensor
+field that describes the space and takes into account all of the specified
+constraints. This is done using an exponential decay combinator, such that when
+a basis is combined with others, the constraint that defined it has less
+influence on the direction of the field the further a point is from it. The
+equations in this section follow closely the system detailed in \cite{chen}.
+
+\paragraph{Linear Constraints}
+A linear constraint produces the grid pattern that is a common basis for many
+cities (consider the streets of Manhattan). The constraint has two parts: a
+magnitude $l$ and a direction $\theta$. The basis field for a linear constraint
+is defined as
+\[
+    T(\pt) =
+        l\sqmat{\cos{2\theta}}{\sin{2\theta}}{\sin{2\theta}}{-\cos{2\theta}}
+\]
+for any point $\bf p$ in the field. It is apparent from this definition that
+the direction of a linear basis field is constant, regardless of the value of
+$\bf p$.
+
+\paragraph{Radial Constraints}
+A radial constraint is used to describe specifically circular areas, such as
+roundabouts or more generally curved roads, common in suburban areas. They are
+defined at a point $(x_0,y_0)$; major streamlines are drawn as concentric circles
+around the point, while minor ones emanate outwards. The basis field for a
+radial constraint at a point $\pt=x_p,y_p$ is defined as
+\[
+    T(\pt) = \sqmat{y^2-x^2}{-2xy}{-2xy}{-(y^2-x^2)}
+\]
+where $x=x_p-x_0$ and $y=x_p-x_0$.
+
+\paragraph{Combining Basis Fields}
+In order to define a tensor field based on multiple basis fields, we use a
+function which sums the basis fields together with an exponential falloff,
+so the tensor field is most strongly influenced by a basis field close to
+the point around which that basis field is defined:
+\[
+    T(\pt) = \sum_i e^{-d||\pt-\pt_i||^2}T_i(\pt)
+\]
+where $d$ is a decay constant, $\pt_i$ is the point at which the $i$th
+constraint is defined, and $T_i$ is the basis field corresponding to the $i$th
+constraint.
+
+\paragraph{Eigenvectors}
+The eigenvalues of a tensor $t$ represented by the matrix $\sqmat{a}{b}{b}{-a}$
+are:
+\[
+    \lambda = \pm\sqrt{a^2+b^2} \\
+\]
+The major eigenvalue, $\lambda_1$, is the larger value, and the minor
+eigenvalue, $\lambda_2$ is the smaller value. The corresponding eigenvectors
+are then
+\[
+    \epsilon_i = \left\{1,\frac{\lambda_i-a}{b}\right\}
+\]
+The eigenvectors are defined if the tensor field is not degenerate (that is,
+$a$ and $b$ are both zero) at that point.
+
+\subsection{Streamline Tracing}
+Mathematically, the basic streamline tracing algorithm is fairly simple,
+however, in order to produce a set of streamlines that approximate a road map,
+there are several important cases to consider. In order to trace the
+streamlines, we are forced to use a heuristic, since it isn’t possible to trace
+a continuous line over a constantly-changing tensor field. In order to
+approximate this, we choose a step size $d_\textrm{step}$. Beginning at a seed
+point, which is chosen using the methods detailed in section XX, we calculate
+the eigenvector at that point, extend a line in that direction with length
+$d_\textrm{step}$, and check if we have reached an end condition.  This process
+is repeated until the line until one of three end conditions is reached:
+\begin{enumerate}
+    \item The point is outside the bounds of the map area
+    \item The point is within a certain distance of an existing streamline.
+    \item The point is part of a cycle
+\end{enumerate}
+
+The first is self-explanatory; if the streamline leaves the boundary of the
+map, then we don’t need to draw it any more. The second is one of the
+alterations we used, based on the work of \cite{chen}. When designing a set of
+roads, it is rare that two streets will approach each other and not connect,
+whether they approach at a larger angle (an intersection) or are running mostly
+parallel to each other (a merge zone). To approximate this, at each point we
+find the nearest neighbor out of the streamlines that have already been drawn.
+If it is within a certain separation distance, we draw a line directly between
+the two points.
+
+\todo{image of that nice circular constraint that worked really well with this}
+
+The third condition arises mainly with radial constraints. Although a true
+hyperstreamline would form a perfect circle around a point, our step-size
+approximation means that the trace is more likely to form a spiral. This is
+obviously a behavior that would not appear in normal maps (other than an
+apocalyptic Boston, perhaps). So, when we detect a streamline circling in on
+itself, we close the gap and end the spiral.
+
+\todo{death spiral and improved version}
+
+\subsection{Seed Placement}
+To begin constructing a streamline, one must choose a seed point—an initial
+coordinate in 2D space from which the streamline is traced. In practice, the
+matter of carefully choosing seed points is very important in generating an
+evenly spaced configuration of roads. We investigated three methods of placing
+seed points:
+\begin{enumerate}
+    \item Random
+    \item Furthest-Point
+    \item Even Streamline Spacing
+\end{enumerate}
+Purely random placement of seed points generally results in both clusters of
+roads which are too close together and large gaps between roads. In the
+furthest-point placement, we place each new seed point within the bounds of the
+space such that the distance to all existing streamlines is maximized. As a
+simplifying heuristic for this method, we generate a number of random candidate
+seed points which is small compared to the infinite number of possible seed
+points (constrained only by floating point precision) and choose the one
+furthest from all existing streamlines. In practice, this offers a significant
+improvement over the random seed placement, and results in a more even
+distribution of streamlines.
+
+However, even choosing well-spaced seed points will not necessarily result in
+an even distribution of streamlines, since two well-spaced seed points can
+result in two streamlines which quickly converge. Therefore, we introduce a
+third method. To assess how evenly spaced a set of streamlines is, we will
+generate a set of discrete points evenly spaced along each streamline, place
+these points into $n\times n$ equal-sized buckets covering the bounds of the
+space, and perform a chi-squared goodness of fit test on distribution of points
+across bucket, where the expected distribution is uniform:
+
+\todo{chi-squared equation}
+
+We will choose the seed point whose corresponding streamline, when added to the
+existing set of streamlines, optimizes the evenness of streamline placement,
+according to the above metric. As with the previous method, we employ a
+simplifying heuristic, generating a relatively small set of candidate seed
+points, tracing a streamline from each, determining which streamline results in
+a more even field of streamlines, and adding it.
+
+\todo{add practical results of this method if we can get it working}
+
 \section{Assessment of our Work}
@@
 
+\onecolumn
 \bibliographystyle{plain}

### 2013-12-12 paper/paper.tex
added chi^2 formula and more explanation, added more latex processing files to gitignore
added assessment section from google doc

@@
 
-\section{Implementation Details}
-
 \subsection{User-Generated Constraints}
+We begin with a JSON parser, which accepts linear and radial constraints.
+Constraints must be input with their type, linear or radial, in addition to
+their position in the domain (in the form of x and y coordinates), and finally,
+if the constraint is linear, it’s direction in radians and magnitude. In
+addition, the executable will accept width and height parameters representing
+the size of the domain to build street maps on. Together, the JSON constraints
+and width/height parameters form the initial input data that will next be used
+to form tensor fields.
 
@@
 
-\paragraph{Eigenvectors}
+\subsection{Eigenvectors}
 The eigenvalues of a tensor $t$ represented by the matrix $\sqmat{a}{b}{b}{-a}$
@@
 across bucket, where the expected distribution is uniform:
-
-\todo{chi-squared equation}
+\[
+    \chi^2 = \sum_{i=1}^{n^2} \frac{(O_i-E_i)^2}{E_i}
+\]
+where $n$ is the number of buckets along each axis, $O_i$ is the observed
+number of points in the $i$th bucket, and $E_i=\frac{m}{n^2}$ is the
+expected number of points in each bucket, assuming a uniform distribution,
+where $m$ is the total number of points in the space.
 
@@
 
-\section{Assessment of our Work}
+\section{Assessment}
+There were two main goals that we set out to achieve with this project. Here we
+present a mostly heuristic, slightly quantitative attempt to qualify our
+success, using both internal evaluation and comparisons to existing systems
+\cite{chen}.
+
+\subsection{Modularity \& Extensibility}
+We believe that we were successful in creating a modular system with reusable
+parts that is easy to understand, extend, and modify. One way to illustrate
+this is by looking at our pipeline. As a refresher, look at Fig. ***. Each
+step of the pipeline has a clear correspondence in the modules that we have
+written. We have two modules to deal with constraints, \texttt{Constraint}. To transform
+constraints to maps, we have written \texttt{TensorField} and \texttt{Tensor} modules, the latter
+of which performs operations on tensors including eigenvector calculations, as
+well as \texttt{Streamline}, which has the ability to calculate seed points and trace
+lines.
+
+We also created several generally useful modules, including \texttt{Vector2},
+which performs basic operations on 2-dimensional vectors and
+\texttt{NearestNeighbor}, which defines a type class for a structure that can
+store \texttt{Vector2}s and query for the nearest neighbor, as well as a
+bucket-based implementation of such a data structure. To deal with input and
+output, we created a \texttt{JSONParser} so that our input can come from any
+source that is capable of producing JSON, and an \texttt{SVGWriter}, which
+deals with the widely-used file format that produces the output seen in many of
+the figures in this paper.
+
+In an attempt to compare this with the original work, we consulted the header
+files available with Chen’s code. The system has significantly more
+functionality than ours, but we had some difficulty discerning which modules
+were parallel to ours. The \texttt{tensoranalysis} contains methods for tensor
+line tracing, and \texttt{ClosedStreamlineTracing} seems to be related,
+although we did not find documentation on the difference between open and
+closed tracing. \texttt{SeedPointCtrlDlg} is a dialogue probably related to the
+OpenGL framework, but didn’t have the actual calculations of seed points. At
+the least, our implementation has a more direct pipeline-to-module correlation,
+which will be beneficial for future contributors who wish to build off of our
+work.
+
+\subsection{Natural Expression of Underlying Mathematics}
+Overall, this seemed to be successful. In general, functional languages are
+well suited to express mathematics. Our decision to implement this program in
+Haskell meant that that we were able to take advantage of this natural
+proclivity towards math in order to make our code easy to understand.
 
@@
 
-\onecolumn
 \bibliographystyle{plain}

### 2013-12-12 paper/paper.tex
pipeline intro

@@
 \section{Pipeline}
+We begin with a set of user constraints, which illustrate the general shape of
+the roads on the blank canvas. These can be linear or radial, and can be input
+as JSON. From these we generate a tensor field, and from that, major and minor
+eigenvector fields, which describe the direction stipulate by the constraints at
+every point in the area. Finally, based on these eigenvector fields, we generate
+seed points from which to trace streamlines across the map. These streamlines
+produce the final roads.
+
+\todo{insert pipeline diagram}
 

### 2013-12-12 paper/paper.tex
added introduction and goals from google doc

--- 2013-12-12 paper/paper.tex
@@
 \newcommand{\todo}[1]{\begin{center}\fbox{\parbox{150pt}{#1}}\end{center}}
+\def \cpp {C\nolinebreak[4]\hspace{-.05em}\raisebox{.4ex}{\tiny\bf ++}~}
 
@@
 \section{Introduction}
+Visual artists often face the time-consuming task of generating extensive,
+realistic scenes by hand for movies and video games. However, the apparent
+randomness found in systems like street layouts can be modelled effectively
+with a variety of mathematical algorithms, giving rise to the growing field of
+procedural generation. In this paper we will discuss an algorithm using tensor
+fields, from [Chen], to generate realistic city street maps. To facilitate the
+generation of aesthetically pleasing maps which fit an artist's requirements,
+we implement a simple system of user-specified constraints to allow the user to
+shape the resulting street map to fit their needs.
 
 \section{Existing Work}
+\textit{Interactive Procedural Street Modelling} \cite{chen} introduces a novel
+new concept of using tensor fields to model realistic street layouts. They 
+offer
+a well-defined pipeline of mathematical operations to go from simple user
+inputs to a final street graph. They also implement a number of operations to
+modify the street graph more granularly once it has been generated, such as
+removal of roads and retracing of the roads in a defined region with a new
+tensor field. The mathematical concepts fundamental to the procedural
+generation algorithm are covered thoroughly and clearly.
+
+While all of the information presented in the paper to be very well-formulated
+and informative, we found the provided implementation of the described
+algorithm more difficult to grasp. The large quantity of \cpp code necessary to
+implement the full-featured interactive system provided by Chen et al. made it
+difficult for us to thoroughly understand the constituent components such that
+we would be able to modify or extend the existing code.
 
 \section{Our Goals}
+We thought we could use the power and expressiveness of a functional language 
+to produce a more modular, extensible, and readable implementation of the core 
+of the algorithm described in Chen et al. which would provide a more 
+accessible starting point for later work to build off of. While we sacrifice 
+the performance offered by the \cpp implementation, which is closely tied to
+OpenGL, we contend that the gains in programmer productivity will more than
+make up for this deficit. While a production implementation of such a pipeline
+should certainly be implemented with efficiency in mind, we believe that, in
+the process of investigating new algorithms, the ease with which the
+implementor can translate an idea or mathematical concept into proof-of-concept
+code is of the utmost importance. Later in the paper, we will assess our
+success in achieving these goals.
 
@@
 Constraints must be input with their type, linear or radial, in addition to
-their position in the domain (in the form of x and y coordinates), and finally,
-if the constraint is linear, it’s direction in radians and magnitude. In
-addition, the executable will accept width and height parameters representing
-the size of the domain to build street maps on. Together, the JSON constraints
-and width/height parameters form the initial input data that will next be used
-to form tensor fields.
+their position in the domain (in the form of $x$ and $y$ coordinates), and
+finally, if the constraint is linear, it’s direction in radians and magnitude.
+In addition, the executable will accept width and height parameters
+representing the size of the domain to build street maps on. Together, the JSON
+constraints and width/height parameters form the initial input data that will
+next be used to form tensor fields.
 
@@
 roundabouts or more generally curved roads, common in suburban areas. They are
-defined at a point $(x_0,y_0)$; major streamlines are drawn as concentric circles
+defined at a point $(x_0,y_0)$; major streamlines are drawn as concentric 
+circles
 around the point, while minor ones emanate outwards. The basis field for a
@@
 step of the pipeline has a clear correspondence in the modules that we have
-written. We have two modules to deal with constraints, \texttt{Constraint}. To transform
-constraints to maps, we have written \texttt{TensorField} and \texttt{Tensor} modules, the latter
+written. We have two modules to deal with constraints, \texttt{Constraint}. To 
+transform
+constraints to maps, we have written \texttt{TensorField} and \texttt{Tensor} 
+modules, the latter
 of which performs operations on tensors including eigenvector calculations, as
-well as \texttt{Streamline}, which has the ability to calculate seed points and trace
+well as \texttt{Streamline}, which has the ability to calculate seed points 
+and trace
 lines.

### 2013-12-12 paper/paper.tex
merge

@@
 inputs to a final street graph. They also implement a number of operations to
-modify the street graph more granularly once it has been generated, such as
-removal of roads and retracing of the roads in a defined region with a new
+modify the street graph more granularly once it has been generated, such as the
+removal and retracing of the roads in a defined region with a new
 tensor field. The mathematical concepts fundamental to the procedural
@@
 
-While all of the information presented in the paper to be very well-formulated
-and informative, we found the provided implementation of the described
+While all of the information presented in the paper completely describes all of
+the necessary mathematics in a clear manner,
+we found the provided implementation of the described
 algorithm more difficult to grasp. The large quantity of \cpp code necessary to
@@
 \]
-A tensor field $T$ is a function that maps from 2D points to tensors. In the
+A tensor field $T$ describes the flow of the constraints across a field,
+and is represented as a function that maps from 2D points to tensors. In the
 context of this project, a tensor is simply a tool we use to determine the

### 2013-12-12 paper/paper.tex

@@
 \begin{abstract}
+This paper addresses the issue of creating modular, extensible code given an
+existing solution to a problem, namely, interactively modeling large street
+networks. We describe an implementation in Haskell of a framework that allows 
+users to create a street network from scratch, based on the techniques described 
+in [Chen]. In particular, we were able to express the underlying mathematics
+more intuitively that is possible in an imperative language.
+
 \end{abstract}

### 2013-12-12 paper/paper.tex
fix grammar in abstract

--- 2013-12-12 paper/paper.tex
@@
 \begin{abstract}
+This paper addresses the issue of creating modular, extensible code given an
+existing solution to a problem, namely, interactively modeling large street
+networks. We describe a functional implementation of a framework that allows
+users to create a realistic street network over an empty domain, based on the
+techniques described
+in [Chen]. In particular, we were able to express the underlying mathematics
+more intuitively than is possible in an imperative language.
+
 \end{abstract}

### 2013-12-12 paper/paper.tex
fix grammar in abstract
made citation style a bit more readable

@@
 \usepackage{amsmath}
+% \usepackage{natbib}
 
@@
 in [Chen]. In particular, we were able to express the underlying mathematics
-more intuitively than is possible in an imperative language.
+more intuitively and clearly than is possible in an imperative language.
 
@@
 procedural generation. In this paper we will discuss an algorithm using tensor
-fields, from [Chen], to generate realistic city street maps. To facilitate the
-generation of aesthetically pleasing maps which fit an artist's requirements,
-we implement a simple system of user-specified constraints to allow the user to
-shape the resulting street map to fit their needs.
+fields, from \cite{Chen}, to generate realistic city street maps. To facilitate
+the generation of aesthetically pleasing maps which fit an artist's
+requirements, we implement a simple system of user-specified constraints to
+allow the user to shape the resulting street map to fit their needs.
 
@@
 new concept of using tensor fields to model realistic street layouts. They
-offer
-a well-defined pipeline of mathematical operations to go from simple user
+offer a well-defined pipeline of mathematical operations to go from simple user
 inputs to a final street graph. They also implement a number of operations to
@@
 \section{Mathematical Background}
-A tensor $t$ is a *magical thing* represented by a 2x2 symmetric, traceless
+A tensor $t$ is a geometric object represented by a 2x2 symmetric, traceless
 matrix:
@@
 influence on the direction of the field the further a point is from it. The
-equations in this section follow closely the system detailed in \cite{chen}.
+equations in this section follow closely the system detailed in \cite{Chen}.
 
@@
 map, then we don’t need to draw it any more. The second is one of the
-alterations we used, based on the work of \cite{chen}. When designing a set of
+alterations we used, based on the work of \cite{Chen}. When designing a set of
 roads, it is rare that two streets will approach each other and not connect,
@@
 success, using both internal evaluation and comparisons to existing systems
-\cite{chen}.
+\cite{Chen}.
 
@@
 
-\bibliographystyle{plain}
+\bibliographystyle{alpha}
 \bibliography{references}

### 2013-12-12 paper/paper.tex
fixed citation capitalization

@@
 \usepackage{amsmath}
-% \usepackage{natbib}
 
@@
 users to create a realistic street network over an empty domain, based on the
-techniques described
-in [Chen]. In particular, we were able to express the underlying mathematics
+techniques described in \textit{Interactive Procedural Street Modelling}
+\cite{chen}. In particular, we were able to express the underlying mathematics
 more intuitively and clearly than is possible in an imperative language.
-
 \end{abstract}
@@
 procedural generation. In this paper we will discuss an algorithm using tensor
-fields, from \cite{Chen}, to generate realistic city street maps. To facilitate
+fields, from \cite{chen}, to generate realistic city street maps. To facilitate
 the generation of aesthetically pleasing maps which fit an artist's
@@
 influence on the direction of the field the further a point is from it. The
-equations in this section follow closely the system detailed in \cite{Chen}.
+equations in this section follow closely the system detailed in \cite{chen}.
 
@@
 map, then we don’t need to draw it any more. The second is one of the
-alterations we used, based on the work of \cite{Chen}. When designing a set of
+alterations we used, based on the work of \cite{chen}. When designing a set of
 roads, it is rare that two streets will approach each other and not connect,
@@
 success, using both internal evaluation and comparisons to existing systems
-\cite{Chen}.
+\cite{chen}.
 

### 2013-12-12 paper/paper.tex
references and figs in underlying math

--- 2013-12-12 paper/paper.tex
@@
 \usepackage{amsmath}
+\usepackage{graphicx}
+
+%borders
+\usepackage{float}
+\floatstyle{boxed}
+\restylefloat{figure}
 
@@
 \section{Mathematical Background}
+\label{sec:math}
 A tensor $t$ is a *magical thing* represented by a 2x2 symmetric, traceless
@@
 
+\begin{figure*}[t!]
+\begin{verbatim}
+-- returns a tuple of the (larger, smaller) eigenvectors of a tensor
+eigenvalues :: Tensor -> (Float, Float)
+eigenvalues (Tensor a b d) =
+  let eval = sqrt (a*a + b*b)
+  in  (eval, (-1)*eval)
+
+-- returns a tuple of the (major, minor) eigenvectors for a tensor
+eigenvectors :: Tensor -> (V2.Vector2, V2.Vector2)
+eigenvectors (Tensor a 0.0 d) =
+  if a > d then (V2.Vector2 1 0, V2.Vector2 0 1)
+     else (V2.Vector2 0 1, V2.Vector2 1 0)
+     eigenvectors (Tensor a b d) =
+       let (eval1, eval2) = eigenvalues (Tensor a b d)
+           mkEvec eval    = V2.Vector2 1 ((eval - a) / b)
+       in (mkEvec eval1, mkEvec eval2)
+\end{verbatim}
+\caption{Haskell implementation of eigenvector calculation}
+\label{fig:evecs}
+\end{figure*}
+
 \subsection{Modularity \& Extensibility}
@@
 
+An example of this can be seen in the calculation of eigenvectors. We derived
+the formula for the eigenvector of a symmetric, traceless tensor 
+(see Section~\ref{sec:math}) and implemented it in the functions in
+Figure~\ref{fig:evecs}.
+
+In [Chen], there is a 75-line function,
+\texttt{cal\_eigenvecs\_onevert\_quad}, to do approximately the same thing. There
+is a fair amount of documentation in the body of the function, explaining what
+each line does, but there also seem to be some side effects relating to global
+variables and large data structures that the authors found difficult to follow,
+mostly by the nature of the language.  Generally speaking, we found that the
+mathematics was significantly easier to express and review in a Haskell
+implementation.
+
 \section{Contributions}

### 2013-12-12 paper/paper.tex

@@
 procedural generation. In this paper we will discuss an algorithm using tensor
-fields, from [Chen], to generate realistic city street maps. To facilitate the
-generation of aesthetically pleasing maps which fit an artist's requirements,
-we implement a simple system of user-specified constraints to allow the user to
-shape the resulting street map to fit their needs.
+fields, from \cite{chen}, to generate realistic city street maps. To facilitate
+the generation of aesthetically pleasing maps which fit an artist's
+requirements, we implement a simple system of user-specified constraints to
+allow the user to shape the resulting street map to fit their needs.
 
@@
 \label{sec:math}
-A tensor $t$ is a *magical thing* represented by a 2x2 symmetric, traceless
+A tensor $t$ is a geometric object represented by a 2x2 symmetric, traceless
 matrix:

### 2013-12-12 paper/paper.tex
add more json files for images

@@
 
-\section{Contributions}
+\section{Conclusion}
+In this paper we have presented the underlying mathematics and functional
+implementation of the work described in \cite{chen}, with the goals being that a
+functional implementation would easily express the mathematics and provide
+modularity that the imperative implementation provide in \cite{chen} could not.
 

### 2013-12-12 Streamline.hs

@@ -108,4 +108,4 @@ placeStreamlines tf nn n fw fh Improved =
 
 traceLines :: TensorField -> Float -> Float -> [SVGElem]
 traceLines tf fw fh =
-  map drawStreamline $ placeStreamlines tf (NN.new fw fh 10) 8 fw fh Random
+  map drawStreamline $ placeStreamlines tf (NN.new fw fh 25) 8 fw fh Furthest

### 2013-12-12 TensorField.hs

@@ -82,7 +82,7 @@ traceStreamline vf nn0 w h p0 step len =
                         Nothing             -> False
             p'      = V2.add p (V2.scalarTimes step (V2.unit dir'))
             output | not inBound = mkStep (p':p:ps) dir' 0
-                   | cycle       = mkStep (p0:p:ps) vlast 0
+                   -- | cycle       = mkStep (p0:p:ps) vlast 0
                    | iSect       = mkStep ((fst $ fromJust lookup):p:ps) dir' 0
                    | otherwise   = mkStep (p':p:ps) dir' (n - 1)
           in output

### 2013-12-12 paper/paper.tex
improved cycle detection figure

@@
 \usepackage{graphicx}
+\usepackage{subcaption}
 
@@
 
-\todo{insert pipeline diagram}
+\begin{figure*}[t!]
+\todo{reimplement pipeline image in \LaTeX}
+\begin{center}
+    \includegraphics[width=5.5in]{images/pipeline.pdf}
+\end{center}
+\caption{Overview of our pipeline}
+\label{fig:pipeline}
+\end{figure*}
 
@@
 The eigenvectors are defined if the tensor field is not degenerate (that is,
-$a$ and $b$ are both zero) at that point.
+$a$ and $b$ are both zero) at that point \cite{find-evs}.
 
@@
 
+\begin{figure*}[t!]
     \todo{image of that nice circular constraint that worked really well with this}
+\caption{Haskell implementation of eigenvector calculation}
+% \label{fig:cycles} name?
+\end{figure*}
 
@@
 
-\todo{death spiral and improved version}
+\floatstyle{plain}
+\restylefloat{figure}
+\begin{figure*}[t!]
+    \centering
+    \begin{subfigure}[b]{2.4in}
+        \fbox{\includegraphics[width=\textwidth]{images/cycle_nodetection.pdf}}
+        \caption{without cycle detection}
+    \end{subfigure}
+    \hspace{20pt}
+    \begin{subfigure}[b]{2.4in}
+        \fbox{\includegraphics[width=\textwidth]{images/cycle_detection.pdf}}
+        \caption{with cycle detection}
+    \end{subfigure}
+\caption{Streamline tracing cycle detection}
+\label{fig:cycles}
+\end{figure*}
+\floatstyle{boxed}
+\restylefloat{figure}
 
@@
 a more even field of streamlines, and adding it.
-
-\todo{add practical results of this method if we can get it working}
 

### 2013-12-12 paper/paper.tex
fixed some references and typos in paper

@@
 \]
+
 The minor eigenvectors are
@@
 \begin{figure*}[t!]
-\todo{reimplement pipeline image in \LaTeX}
 \begin{center}
@@
 next be used to form tensor fields.
+
+% \floatstyle{plain}
+% \restylefloat{figure}
+\begin{figure*}[t!]
+    \centering
+    \begin{subfigure}[b]{2.4in}
+        \begin{verbatim}
+            [{
+              "tycon" : "Linear",
+              "posx" : 2,
+              "posy" : 2,
+              "dir" : 0.01,
+              "mag" : 3
+            },
+            {
+              "tycon" : "Radial",
+              "posx" : 10,
+              "posy" : 10,
+              "dir" : 0,
+              "mag" : 0
+            }]
+        \end{verbatim}
+        \caption{JSON file from by user}
+    \end{subfigure}
+    \hspace{20pt}
+    \begin{subfigure}[b]{2.4in}
+        \todo{image from constraints to left}
+        \caption{generated constraints}
+    \end{subfigure}
+\caption{JSON parsing of constrains}
+\label{fig:json}
+\end{figure*}
+% \floatstyle{boxed}
+% \restylefloat{figure}
 
@@
 \]
-where $x=x_p-x_0$ and $y=x_p-x_0$.
+where $x=x_p-x_0$ and $y=y_p-y_0$.
 
@@
 approximate this, we choose a step size $d_\textrm{step}$. Beginning at a seed
-point, which is chosen using the methods detailed in section XX, we calculate
-the eigenvector at that point, extend a line in that direction with length
-$d_\textrm{step}$, and check if we have reached an end condition.  This process
-is repeated until the line until one of three end conditions is reached:
+point, which is chosen using the methods detailed in Section \ref{seedpoints},
+we calculate the eigenvector at that point, extend a line in that direction
+with length $d_\textrm{step}$, and check if we have reached an end condition.
+This process is repeated until the line until one of three end conditions is
+reached:
 \begin{enumerate}
@@
 \begin{figure*}[t!]
-    \todo{image of that nice circular constraint that worked really well with this}
-\caption{Haskell implementation of eigenvector calculation}
+    \todo{image of that nice circular constraint that worked really well with
+    this}
 % \label{fig:cycles} name?
@@
 
-\subsection{Seed Placement}
+\subsection{Seed Placement}\label{seedpoints}
 To begin constructing a streamline, one must choose a seed point—an initial
@@
 \end{verbatim}
-\caption{Haskell implementation of eigenvector calculation}
+\caption{Haskell implementation of eigenvector calculation (from
+            \texttt{Tensor.hs})}
 \label{fig:evecs}
@@
 parts that is easy to understand, extend, and modify. One way to illustrate
-this is by looking at our pipeline. As a refresher, look at Fig. ***. Each
-step of the pipeline has a clear correspondence in the modules that we have
-written. We have two modules to deal with constraints, \texttt{Constraint}. To
-transform
+this is by looking at our pipeline (Figure \ref{fig:pipeline}). Each step of
+the pipeline has a clear correspondence in the modules that we have written.
+We have two modules to deal with constraints, \texttt{Constraint}. To transform
 constraints to maps, we have written \texttt{TensorField} and \texttt{Tensor}
-modules, the latter
-of which performs operations on tensors including eigenvector calculations, as
-well as \texttt{Streamline}, which has the ability to calculate seed points
-and trace
-lines.
+modules, the latter of which performs operations on tensors including
+eigenvector calculations, as well as \texttt{Streamline}, which has the ability
+to calculate seed points and trace lines.
 

### 2013-12-12 Main.hs
add seed placement parsing as input argument

@@ -4,35 +4,45 @@ import SVGWriter
 import JSONParser
 import Control.Applicative
 import Data.Aeson
+import Streamline (PlacementMethod (Furthest, Random, Improved))
 import qualified Streamline as SL
 import System.Environment
 import System.Exit
 import qualified Data.ByteString.Lazy as B
 
-makeSVG :: [Constraint] -> Float -> Float -> IO ()
-makeSVG cs fw fh =
+makeSVG :: [Constraint] -> Float -> Float -> PlacementMethod -> IO ()
+makeSVG cs fw fh method =
     let tf = TF.makeTensorField cs
     in putStrLn $ writeSVG (appendElements
                            (TF.plotTensorField tf cs fw fh)
-                           (SL.traceLines tf fw fh))
+                           (SL.traceLines tf fw fh method))
 
 main :: IO ()
 main = getArgs >>= parseCmdArgs
 
-parseCmdArgs ["-h"]              = usage                               >> exit
-parseCmdArgs ["-v"]              = version                             >> exit
-parseCmdArgs [file, fw, fh]      = (buildViz file (read fw) (read fh)) >> exit
-parseCmdArgs [file, fw, fh, opt] = (buildViz file (read fw) (read fh)) >> exit
-parseCmdArgs _                   = usage                               >> exit
+def :: String
+def = "furthest"
+
+parseCmdArgs ["-h"]            = usage                                   >> exit
+parseCmdArgs ["-v"]            = version                                 >> exit
+parseCmdArgs [file, fw, fh]    = (buildViz file (read fw) (read fh) def) >> exit
+parseCmdArgs [file, fw, fh, o] = (buildViz file (read fw) (read fh) o)   >> exit
+parseCmdArgs _                 = usage                                   >> exit
 
 usage   = putStrLn "usage: tensor <inputFile> <fieldWidth> <fieldHeight> [opt]"
 version = putStrLn "Tensor v0.1"
 exit    = exitWith ExitSuccess
 die     = exitWith (ExitFailure 1)
 
-buildViz :: FilePath -> Float -> Float -> IO ()
-buildViz inputFile fw fh = do
+buildViz :: FilePath -> Float -> Float -> String -> IO ()
+buildViz inputFile fw fh pmeth = do
     d <- (eitherDecode <$> (B.readFile inputFile) :: IO (Either String [Input]))
     case d of
       Left err    -> putStrLn $ "Failure on input: " ++ err
[…]

### 2013-12-12 Streamline.hs

@@ -1,4 +1,5 @@
-module Streamline (PlacementMethod, traceLines) where
+module Streamline (PlacementMethod (Random, Furthest, Improved), traceLines)
+where
 
 import TensorField as TF
 import Vector2 (Vector2 (Vector2))
@@ -106,6 +107,6 @@ placeStreamlines tf nn n fw fh Furthest =
 placeStreamlines tf nn n fw fh Improved =
   fst $ placeSeedsImproved (tensorfieldEigenvectors tf) nn [] n fw fh
 
-traceLines :: TensorField -> Float -> Float -> [SVGElem]
-traceLines tf fw fh =
-  map drawStreamline $ placeStreamlines tf (NN.new fw fh 25) 8 fw fh Furthest
+traceLines :: TensorField -> Float -> Float -> PlacementMethod -> [SVGElem]
+traceLines tf fw fh method =
+  map drawStreamline $ placeStreamlines tf (NN.new fw fh 25) 8 fw fh method

### 2013-12-12 Main.hs
add error message

@@ -43,6 +43,7 @@ buildViz inputFile fw fh pmeth = do
                                     "furthest" -> Furthest
                                     "random"   -> Random
                                     "improved" -> Improved
+                                    _          -> error "Seed placement method not found."
                      in makeSVG (map inputToConstraint input) fw fh method
 
 

### 2013-12-12 paper/paper.tex
made json input figure more readable
removed some figure outlines

@@
 \usepackage{subcaption}
-
-%borders
-\usepackage{float}
-\floatstyle{boxed}
-\restylefloat{figure}
+\usepackage{fancyvrb}
 
@@
 This paper addresses the issue of creating modular, extensible code given an
-existing solution to a problem, namely, interactively modeling large street
+existing solution to a problem, namely, procedurally modeling large street
 networks. We describe a functional implementation of a framework that allows
@@
 the generation of aesthetically pleasing maps which fit an artist's
-requirements, we implement a simple system of user-specified constraints to
-allow the user to shape the resulting street map to fit their needs.
+requirements, we implement a simple system by which the user can input
+constraints to the algorithm, shaping the resulting street map to fit their
+needs.
 
@@
 While all of the information presented in the paper completely describes all of
-the necessary mathematics in a clear manner,
-we found the provided implementation of the described
-algorithm more difficult to grasp. The large quantity of \cpp code necessary to
-implement the full-featured interactive system provided by Chen et al. made it
-difficult for us to thoroughly understand the constituent components such that
-we would be able to modify or extend the existing code.
+the necessary mathematics for the algorithm in a clear manner, we found the
+provided implementation more difficult to grasp. The large quantity of \cpp
+code necessary to implement the full-featured interactive system provided by
+Chen et al. made it difficult for us to thoroughly understand the constituent
+components such that we would be able to modify or extend the existing code.
 
 \section{Our Goals}
-We thought we could use the power and expressiveness of a functional language
-to produce a more modular, extensible, and readable implementation of the core
-of the algorithm described in Chen et al. which would provide a more
-accessible starting point for later work to build off of. While we sacrifice
-the performance offered by the \cpp implementation, which is closely tied to
-OpenGL, we contend that the gains in programmer productivity will more than
-make up for this deficit. While a production implementation of such a pipeline
-should certainly be implemented with efficiency in mind, we believe that, in
-the process of investigating new algorithms, the ease with which the
-implementor can translate an idea or mathematical concept into proof-of-concept
-code is of the utmost importance. Later in the paper, we will assess our
-success in achieving these goals.
+We approached the problem with the idea that we could use the power and
+expressiveness of a functional language to produce a more modular, extensible,
+and readable implementation of the core of the algorithm described in Chen et
+al. which would provide a more accessible starting point for later work to
+build off of. While we sacrifice the performance offered by the \cpp
+implementation, which is closely tied to OpenGL, we contend that the gains in
+programmer productivity will alleviate this deficit. While a production
+implementation of such a pipeline should certainly be implemented with
+efficiency in mind, we believe that, in the process of investigating new
+algorithms, the ease with which the implementor can translate an idea or
+mathematical concept into proof-of-concept code is of the utmost importance.
+Later in the paper, we will assess our success in achieving these goals.
 
-\section{Mathematical Background}
-\label{sec:math}
-A tensor $t$ is a geometric object represented by a 2x2 symmetric, traceless
-matrix:
+\section{Mathematical Background}\label{sec:math}
+In this section we will introduce the mathematical concepts necessary to
+understand our algorithm. A tensor $t$ is a geometric object represented by a
+2x2 symmetric, traceless matrix:
 \[
@@
 
-The major eigenvectors of $t$ are
+The major eigenvectors of $t$ are:
 \[
@@
 
-The minor eigenvectors are
+The minor eigenvectors are:
 \[
@@
 \]
-A tensor field $T$ describes the flow of the constraints across a field,
-and is represented as a function that maps from 2D points to tensors. In the
-context of this project, a tensor is simply a tool we use to determine the
-direction of a field at a point. This direction is found by calculating the
-eigenvectors of the tensors. Major eigenvectors are parallel to the field;
-minor eigenvectors are perpendicular.
 
-Another important concept is the hyperstreamline. This is an effective way to
-visualize tensor fields. It is created by drawing a line through an eigenvector
-field such that the line is tangent to every eigenvector along its path. In
-this setting, there are major and minor hyperstreamlines that correspond to the
-major and minor eigenvectors. The major and minor eigenvectors of a tensor are
-guaranteed to be perpendicular for any tensor, unless it is a degenerate point
-in the tensor field ($T(\bf p)=0$), for which eigenvectors cannot be
-calculated. This property leads to the generation of perpendicular street
-intersections, a common feature of real-world road networks, especially in
-cities.
+A tensor field $T$ describes the flow of the constraints across a field, and is
+represented as a function that maps from 2D points to tensors. In the context
+of this project, a tensor a tool we use to determine the direction of a field
+at a given point. This direction is found by calculating the eigenvectors of
+the tensors. Major eigenvectors are parallel to the field; minor eigenvectors
+are perpendicular.
+
+Another important concept is the hyperstreamline, an effective way to visualize
+tensor fields. A hyperstreamline is created by drawing a line through an
+eigenvector field such that the line is tangent to every eigenvector along its
+path. In this setting, there are major and minor hyperstreamlines that
+correspond to the major and minor eigenvectors. The major and minor
+eigenvectors of a tensor are guaranteed to be perpendicular for any tensor,
+unless it is a degenerate point in the tensor field ($T(\bf p)=0$), for which
+eigenvectors cannot be calculated. This property leads to the generation of
+perpendicular street intersections, a common feature of real-world road
+networks, especially in cities.
 
@@
 as JSON. From these we generate a tensor field, and from that, major and minor
-eigenvector fields, which describe the direction stipulate by the constraints at
-every point in the area. Finally, based on these eigenvector fields, we generate
-seed points from which to trace streamlines across the map. These streamlines
-produce the final roads.
+eigenvector fields, which describe the direction stipulated by the constraints
+at every point in the area. Finally, based on these eigenvector fields, we
+generate seed points from which streamlines are traced across the map. These
+streamlines produce the final roads.
 
@@
     \centering
+    \vspace{10pt}
     \begin{subfigure}[b]{2.4in}
@@
     \begin{subfigure}[b]{2.4in}
-        \todo{image from constraints to left}
+        \fbox{\includegraphics[width=\textwidth]{images/json.pdf}}
+        \vspace{3pt}
         \caption{generated constraints}
     \end{subfigure}
-\caption{JSON parsing of constrains}
+    \caption{JSON input constraint format}
 \label{fig:json}
 \end{figure*}
-% \floatstyle{boxed}
-% \restylefloat{figure}
 
@@
 magnitude $l$ and a direction $\theta$. The basis field for a linear constraint
-is defined as
+is defined as:
 \[
@@
 defined at a point $(x_0,y_0)$; major streamlines are drawn as concentric
-circles
-around the point, while minor ones emanate outwards. The basis field for a
-radial constraint at a point $\pt=x_p,y_p$ is defined as
+circles around the point, while minor ones emanate outwards. The basis field
+for a radial constraint at a point $\pt=x_p,y_p$ is defined as
 \[
@@
 are:
-\[
-    \lambda = \pm\sqrt{a^2+b^2} \\
-\]
-The major eigenvalue, $\lambda_1$, is the larger value, and the minor
-eigenvalue, $\lambda_2$ is the smaller value. The corresponding eigenvectors
-are then
+\begin{align}
+    \lambda_1 &= \sqrt{a^2+b^2} \\
+    \lambda_2 &= -\lambda_1 \\
+\end{align}
+
+The corresponding eigenvectors are:
 \[
@@
 \]
+
 The eigenvectors are defined if the tensor field is not degenerate (that is,
@@
 with length $d_\textrm{step}$, and check if we have reached an end condition.
-This process is repeated until the line until one of three end conditions is
+This process is repeated on the line until one of three end conditions is
 reached:
@@
 The first is self-explanatory; if the streamline leaves the boundary of the
-map, then we don’t need to draw it any more. The second is one of the
-alterations we used, based on the work of \cite{chen}. When designing a set of
-roads, it is rare that two streets will approach each other and not connect,
-whether they approach at a larger angle (an intersection) or are running mostly
-parallel to each other (a merge zone). To approximate this, at each point we
-find the nearest neighbor out of the streamlines that have already been drawn.
-If it is within a certain separation distance, we draw a line directly between
-the two points.
+map, then we don’t need to draw it any more.
 
-\begin{figure*}[t!]
-    \todo{image of that nice circular constraint that worked really well with
-    this}
-% \label{fig:cycles} name?
-\end{figure*}
+The second is one of the alterations we used, based on the work of \cite{chen}.
+When designing a set of roads, it is rare that two streets will approach each
+other and not connect, whether they approach at a larger angle (an
+intersection) or are running mostly parallel to each other (a merge zone). To
+approximate this, at each point we find the nearest neighbor out of the
+streamlines that have already been drawn.  If it is within a certain separation
+distance, we draw a line directly between the two points.
 
@@
 
-\floatstyle{plain}
-\restylefloat{figure}
+\begin{figure*}[t!]
+    \centering
+    \fbox{\includegraphics[width=4in]{images/linear.pdf}}
+    \caption{Streamline tracing near road merging}
+\end{figure*}
+
 \begin{figure*}[t!]
@@
 \end{figure*}
-\floatstyle{boxed}
-\restylefloat{figure}
 
 \subsection{Seed Placement}\label{seedpoints}
-To begin constructing a streamline, one must choose a seed point—an initial
+To begin constructing a streamline, one must choose a seed point—--an initial
 coordinate in 2D space from which the streamline is traced. In practice, the
@@
     \item Furthest-Point
-    \item Even Streamline Spacing
+    \item Even Streamline Spacing ($\chi^2$ distribution)
 \end{enumerate}
+
 Purely random placement of seed points generally results in both clusters of
-roads which are too close together and large gaps between roads. In the
-furthest-point placement, we place each new seed point within the bounds of the
-space such that the distance to all existing streamlines is maximized. As a
-simplifying heuristic for this method, we generate a number of random candidate
-seed points which is small compared to the infinite number of possible seed
-points (constrained only by floating point precision) and choose the one
-furthest from all existing streamlines. In practice, this offers a significant
-improvement over the random seed placement, and results in a more even
-distribution of streamlines.
+roads which are too close together and large gaps between roads.
+
+In the furthest-point placement, we place each new seed point within the bounds
+of the space such that the distance to all existing streamlines is maximized.
+As a simplifying heuristic for this method, we generate a number of random
+candidate seed points which is small compared to the infinite number of
+possible seed points (constrained only by floating point precision) and choose
+the one furthest from all existing streamlines. In practice, this offers a
+significant improvement over the random seed placement, and results in a more
+even distribution of streamlines.
 
@@
 points, tracing a streamline from each, determining which streamline results in
-a more even field of streamlines, and adding it.
+a more even field of streamlines, and adding it to the existing set of
+streamlines.
 
@@
 \begin{figure*}[t!]
-\begin{verbatim}
+\begin{Verbatim}[frame=single]
 -- returns a tuple of the (larger, smaller) eigenvectors of a tensor
@@
      in (mkEvec eval1, mkEvec eval2)
-\end{verbatim}
+\end{Verbatim}
+
 \caption{Haskell implementation of eigenvector calculation (from

### 2013-12-12 paper/paper.tex
added more references and images for figures

@@
 
-\section{Pipeline}
+\section{Pipeline}\label{sec:pipeline}
 We begin with a set of user constraints, which illustrate the general shape of
@@
 
-\subsection{Seed Placement}\label{seedpoints}
+\subsection{Seed Placement}\label{sec:seedpoints}
 To begin constructing a streamline, one must choose a seed point—--an initial
@@
 \section{Assessment}
-There were two main goals that we set out to achieve with this project. Here we
-present a mostly heuristic, slightly quantitative attempt to qualify our
-success, using both internal evaluation and comparisons to existing systems
-\cite{chen}.
+We set out to achieve two goals: modularity and expressiveness. In this section
+we will examine the extent to which these goals were achieved.
+
+\subsection{Modularity and Extensibility}
+In Section \ref{sec:seedpoints}, we outlined three methods of seedpoint
+placement: random, furthest-point, and $\chi^2$ even spacing. All three
+are implemented by functions in the \texttt{Streamline} module, and follow a
+similar template. Extending this module with other methods is as simple as
+implementing a new function with the same type signature as
+\texttt{placeSeeds}, adding a new value constructor to the datatype
+\texttt{PlacementMethod}, and adding a new case to the function
+\texttt{placeStreamlines} to match the new data constructor. The purity of
+functions in a language such as Haskell makes it simple to drop in replacements
+for existing functions because one does not have to consider possible side
+effects, but only the types of the inputs and output.
+
+In order to detect intersections between streamlines as they are traced, we
+defined the \texttt{NearestNeighbor} module to implement a data structure which
+stored a set of 2D points and could be queried to determine which of the stored
+points was closest to the query point. This is implemented using a type
+family, with an instance utilizing a blocked 2D array implementation. This
+instance was sufficient for testing purposes, but is currently the performance
+bottleneck in streamline tracing. A new instance of the \texttt{Storage}
+type family could easily be defined in accordance with the type signature to
+provide a more efficient representation. For example, a k-d tree instance of
+the type family could offer higher performance. This well-defined abstraction
+boundary makes this aspect of our pipeline very easily extensible.
+
+\subsection{Expressiveness}
+As our work was not a complete reimplementation of the system described in
+\cite{chen}
 
@@
 \subsection{Modularity \& Extensibility}
-We believe that we were successful in creating a modular system with reusable
-parts that is easy to understand, extend, and modify. One way to illustrate
-this is by looking at our pipeline (Figure \ref{fig:pipeline}). Each step of
-the pipeline has a clear correspondence in the modules that we have written.
-We have two modules to deal with constraints, \texttt{Constraint}. To transform
-constraints to maps, we have written \texttt{TensorField} and \texttt{Tensor}
-modules, the latter of which performs operations on tensors including
-eigenvector calculations, as well as \texttt{Streamline}, which has the ability
-to calculate seed points and trace lines.
-
-We also created several generally useful modules, including \texttt{Vector2},
-which performs basic operations on 2-dimensional vectors and
-\texttt{NearestNeighbor}, which defines a type class for a structure that can
-store \texttt{Vector2}s and query for the nearest neighbor, as well as a
-bucket-based implementation of such a data structure. To deal with input and
-output, we created a \texttt{JSONParser} so that our input can come from any
-source that is capable of producing JSON, and an \texttt{SVGWriter}, which
-deals with the widely-used file format that produces the output seen in many of
-the figures in this paper.
 

### 2013-12-12 paper/paper.tex
rewrote assessment and conclusion

@@
 understand our algorithm. A tensor $t$ is a geometric object represented by a
-2x2 symmetric, traceless matrix:
+$2\times2$ symmetric, traceless matrix:
 \[
@@
 
-\subsection{Eigenvectors}
+\subsection{Eigenvectors}\label{sec:eigenvectors}
 The eigenvalues of a tensor $t$ represented by the matrix $\sqmat{a}{b}{b}{-a}$
@@
 approximate this, we choose a step size $d_\textrm{step}$. Beginning at a seed
-point, which is chosen using the methods detailed in Section \ref{seedpoints},
-we calculate the eigenvector at that point, extend a line in that direction
-with length $d_\textrm{step}$, and check if we have reached an end condition.
-This process is repeated on the line until one of three end conditions is
-reached:
+point, which is chosen using the methods detailed in Section
+\ref{sec:seedpoints}, we calculate the eigenvector at that point, extend a line
+in that direction with length $d_\textrm{step}$, and check if we have reached
+an end condition.  This process is repeated on the line until one of three end
+conditions is reached:
+
 \begin{enumerate}
@@
 As our work was not a complete reimplementation of the system described in
-\cite{chen}
+\cite{chen}, it is difficult to compare the relative expressiveness of our
+Haskell code with the \cpp code of the original implementation. However, we
+will present one comparison of functions from our respective implementations
+with significantly overlapping functionality. Our Haskell code for calculating
+the eigenvectors for a tensor can be found in Figure \ref{fig:evecs}. The
+reader may note the close correspondence between the body of these functions
+and the mathematical formulas used for calculating eigenvectors in Section
+\ref{sec:eigenvectors}.  The 75-line function
+\texttt{cal\_eigenvecs\_onevert\_quad} in Chen's implementation. While it is
+documented, we found it difficult to determine the side effects it had. In a
+language without mutable state or side effect, we found it significantly easier
+to fully understand the behavior of functions.  In contrast, even well-written
+code in \cpp or another language in the imperative paradigm can be difficult to
+understand, because the possibility for side effect requires the reader to have
+an understanding of any part of the system that the function can affect, which
+in practice can make the effort to understand any small part of a large
+codebase much greater.
 
@@
 
-\subsection{Modularity \& Extensibility}
-
-In an attempt to compare this with the original work, we consulted the header
-files available with Chen’s code. The system has significantly more
-functionality than ours, but we had some difficulty discerning which modules
-were parallel to ours. The \texttt{tensoranalysis} contains methods for tensor
-line tracing, and \texttt{ClosedStreamlineTracing} seems to be related,
-although we did not find documentation on the difference between open and
-closed tracing. \texttt{SeedPointCtrlDlg} is a dialogue probably related to the
-OpenGL framework, but didn’t have the actual calculations of seed points. At
-the least, our implementation has a more direct pipeline-to-module correlation,
-which will be beneficial for future contributors who wish to build off of our
-work.
-
-\subsection{Natural Expression of Underlying Mathematics}
-Overall, this seemed to be successful. In general, functional languages are
-well suited to express mathematics. Our decision to implement this program in
-Haskell meant that that we were able to take advantage of this natural
-proclivity towards math in order to make our code easy to understand.
-
-An example of this can be seen in the calculation of eigenvectors. We derived
-the formula for the eigenvector of a symmetric, traceless tensor
-(see Section~\ref{sec:math}) and implemented it in the functions in
-Figure~\ref{fig:evecs}.
-
-In [Chen], there is a 75-line function,
-\texttt{cal\_eigenvecs\_onevert\_quad}, to do approximately the same thing. There
-is a fair amount of documentation in the body of the function, explaining what
-each line does, but there also seem to be some side effects relating to global
-variables and large data structures that the authors found difficult to follow,
-mostly by the nature of the language.  Generally speaking, we found that the
-mathematics was significantly easier to express and review in a Haskell
-implementation.
-
 \section{Conclusion}
 In this paper we have presented the underlying mathematics and functional
-implementation of the work described in \cite{chen}, with the goals being that a
-functional implementation would easily express the mathematics and provide
+implementation of the work described in \cite{chen}, with the goals being that
+a functional implementation would readily express the mathematics, provide
 modularity that the imperative implementation provide in \cite{chen} could not.
+We believe that, due to the focus on expressibility and modularity, assessed in
+the previous section, that an individual looking to understand the pipeline
+will find our code undestandable and approachable, and lend itself to be
+easily extended. We hope that our work will motivate a greater use of
+functional languages in future efforts in the research of new procedural
+algorithms.
 

### 2013-12-13 paper/paper.tex
submitted version of paper

@@
 \usepackage{fancyvrb}
+
+\columnsep=2em
 
@@
 \begin{abstract}
-This paper addresses the issue of creating modular, extensible code given an
-existing solution to a problem, namely, procedurally modeling large street
-networks. We describe a functional implementation of a framework that allows
-users to create a realistic street network over an empty domain, based on the
-techniques described in \textit{Interactive Procedural Street Modelling}
+This paper addresses the issue of creating modular, extensible code to
+implement a pipeline for procedurally modeling large street networks. We
+describe a functional implementation of a framework that allows users to create
+a realistic street network over an empty domain, based on the techniques
+described in \textit{Interactive Procedural Street Modelling} by Chen et al.
 \cite{chen}. In particular, we were able to express the underlying mathematics
-more intuitively and clearly than is possible in an imperative language.
+more intuitively and clearly than is possible in an imperative language, and
+produce a more modular system.
 \end{abstract}
@@
 Visual artists often face the time-consuming task of generating extensive,
-realistic scenes by hand for movies and video games. However, the apparent
-randomness found in systems like street layouts can be modelled effectively
-with a variety of mathematical algorithms, giving rise to the growing field of
-procedural generation. In this paper we will discuss an algorithm using tensor
-fields, from \cite{chen}, to generate realistic city street maps. To facilitate
-the generation of aesthetically pleasing maps which fit an artist's
-requirements, we implement a simple system by which the user can input
-constraints to the algorithm, shaping the resulting street map to fit their
-needs.
+realistic scenes for movies and video games. However, the apparent randomness
+found in systems like street layouts can be modelled effectively with a variety
+of mathematical algorithms, giving rise to the growing field of procedural
+generation. In this paper we will discuss an algorithm using tensor fields,
+from \cite{chen}, to generate realistic city street maps. To facilitate the
+generation of aesthetically pleasing maps which fit an artist's requirements,
+we implement a simple system by which the user can input constraints to the
+algorithm, shaping the resulting street map to fit their needs.
 
 \section{Existing Work}
-\textit{Interactive Procedural Street Modelling} \cite{chen} introduces a novel
-new concept of using tensor fields to model realistic street layouts. They
-offer a well-defined pipeline of mathematical operations to go from simple user
-inputs to a final street graph. They also implement a number of operations to
-modify the street graph more granularly once it has been generated, such as the
-removal and retracing of the roads in a defined region with a new
-tensor field. The mathematical concepts fundamental to the procedural
-generation algorithm are covered thoroughly and clearly.
+\textit{Interactive Procedural Street Modelling} \cite{chen} introduces the
+novel concept of using tensor fields to model realistic street layouts. They
+offer a well-defined pipeline of mathematical operations to generate a street
+graph from user input. They also implement a number of operations to modify the
+street graph more granularly once it has been generated, such as the removal
+and retracing of the roads in a defined region with a new tensor field. The
+mathematical concepts fundamental to the procedural generation algorithm are
+covered thoroughly and clearly.
 
-While all of the information presented in the paper completely describes all of
-the necessary mathematics for the algorithm in a clear manner, we found the
-provided implementation more difficult to grasp. The large quantity of \cpp
+While all of the information presented in the paper completely describes the
+necessary mathematics for the algorithm in an understandable manner, we found
+the provided implementation more difficult to grasp. The large quantity of \cpp
 code necessary to implement the full-featured interactive system provided by
-Chen et al. made it difficult for us to thoroughly understand the constituent
+\cite{chen} made it difficult for us to thoroughly understand the constituent
 components such that we would be able to modify or extend the existing code.
@@
 \section{Our Goals}
-We approached the problem with the idea that we could use the power and
-expressiveness of a functional language to produce a more modular, extensible,
-and readable implementation of the core of the algorithm described in Chen et
-al. which would provide a more accessible starting point for later work to
-build off of. While we sacrifice the performance offered by the \cpp
-implementation, which is closely tied to OpenGL, we contend that the gains in
-programmer productivity will alleviate this deficit. While a production
-implementation of such a pipeline should certainly be implemented with
-efficiency in mind, we believe that, in the process of investigating new
-algorithms, the ease with which the implementor can translate an idea or
-mathematical concept into proof-of-concept code is of the utmost importance.
-Later in the paper, we will assess our success in achieving these goals.
+We approached the problem with the idea that we could use the expressive power
+of a functional language to produce a more modular, extensible, and readable
+implementation of the core of the algorithm described in \cite{chen} which
+would provide a more accessible starting point from which later work can be
+built.  While we sacrifice the performance offered by the \cpp implementation,
+which is closely tied to OpenGL, we contend that the gains in programmer
+productivity will alleviate this deficit. While a production implementation of
+such a pipeline should certainly be implemented with efficiency in mind, we
+believe that, in the process of investigating new algorithms, the ease with
+which the implementor can translate an idea or mathematical concept into
+proof-of-concept code is of the utmost importance.  Later in the paper, we will
+evaluate our success in achieving these goals.
 
@@
 represented as a function that maps from 2D points to tensors. In the context
-of this project, a tensor a tool we use to determine the direction of a field
-at a given point. This direction is found by calculating the eigenvectors of
-the tensors. Major eigenvectors are parallel to the field; minor eigenvectors
-are perpendicular.
+of this project, a tensor is a tool we use to determine the direction of a
+field at a given point. This direction is found by calculating the eigenvectors
+of the tensors. Major eigenvectors are parallel to the field; minor
+eigenvectors are perpendicular.
 
@@
 perpendicular street intersections, a common feature of real-world road
-networks, especially in cities.
+networks.
 
-\section{Pipeline}\label{sec:pipeline}
-We begin with a set of user constraints, which illustrate the general shape of
-the roads on the blank canvas. These can be linear or radial, and can be input
-as JSON. From these we generate a tensor field, and from that, major and minor
-eigenvector fields, which describe the direction stipulated by the constraints
-at every point in the area. Finally, based on these eigenvector fields, we
+\section{Pipeline}\label{sec:pipeline} We begin with a set of user constraints,
+which illustrate the general shape of the roads on the blank canvas. These can
+be linear or radial, and can be input as JSON. From these we generate a tensor
+field, and from that, major and minor eigenvector fields (functions mapping
+from a point to the eigenvector of the tensor defined at that point in the
+tensor field), which describe the direction stipulated by the constraints at
+every point in the area. Finally, based on these eigenvector fields, we
 generate seed points from which streamlines are traced across the map. These
-streamlines produce the final roads.
+streamlines are the final roads.
+
+\subsection{User-Generated Constraints}
+We begin with a JSON parser, which accepts user-given constraints. There are
+two basic types of constraints. Linear constraints have a direction and a
+magnitude and are located at a specific point in the plane. Radial constraints
+are defined only by their location. An example of JSON input and the
+set of constraints it is interpreted as can be found in Figure \ref{fig:json}.
 
@@
 
-\subsection{User-Generated Constraints}
-We begin with a JSON parser, which accepts linear and radial constraints.
-Constraints must be input with their type, linear or radial, in addition to
-their position in the domain (in the form of $x$ and $y$ coordinates), and
-finally, if the constraint is linear, it’s direction in radians and magnitude.
-In addition, the executable will accept width and height parameters
-representing the size of the domain to build street maps on. Together, the JSON
-constraints and width/height parameters form the initial input data that will
-next be used to form tensor fields.
-
-% \floatstyle{plain}
-% \restylefloat{figure}
 \begin{figure*}[t!]
@@
         \end{verbatim}
-        \caption{JSON file from by user}
+        \caption{JSON file entered by user}
     \end{subfigure}
@@
 field that describes the space and takes into account all of the specified
-constraints. This is done using an exponential decay combinator, such that when
-a basis is combined with others, the constraint that defined it has less
-influence on the direction of the field the further a point is from it. The
-equations in this section follow closely the system detailed in \cite{chen}.
+constraints. The equations in this section follow closely the system detailed
+in \cite{chen}.
 
@@
 A linear constraint produces the grid pattern that is a common basis for many
-cities (consider the streets of Manhattan). The constraint has two parts: a
+cities---consider the streets of Manhattan. The constraint has two parts: a
 magnitude $l$ and a direction $\theta$. The basis field for a linear constraint
@@
 \paragraph{Radial Constraints}
-A radial constraint is used to describe specifically circular areas, such as
-roundabouts or more generally curved roads, common in suburban areas. They are
-defined at a point $(x_0,y_0)$; major streamlines are drawn as concentric
-circles around the point, while minor ones emanate outwards. The basis field
-for a radial constraint at a point $\pt=x_p,y_p$ is defined as
+A radial constraint is used to produce circular areas, such as roundabouts and
+other curved roads, common in suburban areas. They are defined at a point
+$(x_0,y_0)$ Major streamlines are drawn as concentric circles around the point,
+while minor ones radiate outwards. The basis field for a radial constraint at a
+point $\pt=(x_p,y_p)$ is defined as:
+
 \[
@@
 are:
-\begin{align}
+\begin{align*}
     \lambda_1 &= \sqrt{a^2+b^2} \\
-    \lambda_2 &= -\lambda_1 \\
-\end{align}
+    \lambda_2 &= -\lambda_1
+\end{align*}
 
@@
 
-The eigenvectors are defined if the tensor field is not degenerate (that is,
-$a$ and $b$ are both zero) at that point \cite{find-evs}.
+The eigenvectors are defined if the tensor field is not degenerate (if $a$ and
+$b$ are both zero, the tensor is degenerate) at that point \cite{find-evs}. The
+eigenvector corresponding to $\lambda_1$ is the major eigenvector, and the
+eigenvector corresponding to $\lambda_2$ is the minor eigenvector.
 
@@
 there are several important cases to consider. In order to trace the
-streamlines, we are forced to use a heuristic, since it isn’t possible to trace
+streamlines, we are forced to use a heuristic, since it isn't possible to trace
 a continuous line over a constantly-changing tensor field. In order to
-approximate this, we choose a step size $d_\textrm{step}$. Beginning at a seed
+approximate this, we choose a step size, $d_\textrm{step}$. Beginning at a seed
 point, which is chosen using the methods detailed in Section
@@
 The first is self-explanatory; if the streamline leaves the boundary of the
-map, then we don’t need to draw it any more.
+map, then we don't need to draw it any more.
 
-The second is one of the alterations we used, based on the work of \cite{chen}.
 When designing a set of roads, it is rare that two streets will approach each
 other and not connect, whether they approach at a larger angle (an
-intersection) or are running mostly parallel to each other (a merge zone). To
-approximate this, at each point we find the nearest neighbor out of the
-streamlines that have already been drawn.  If it is within a certain separation
-distance, we draw a line directly between the two points.
+intersection) or are running mostly parallel to each other (a merge zone). The
+second end condition approximates this; at each point we find the closest point
+on any of the streamlines that have already been traced.  If it is within a
+certain separation distance, $d_\textrm{sep}$, we draw a line directly between
+the two points.
 
@@
 approximation means that the trace is more likely to form a spiral. This is
-obviously a behavior that would not appear in normal maps (other than an
-apocalyptic Boston, perhaps). So, when we detect a streamline circling in on
-itself, we close the gap and end the spiral.
+obviously a behavior that would not appear in normal maps (other than Boston,
+perhaps). So, when we detect a streamline circling in on itself, we close the
+gap and end the spiral.
 
@@
 evenly spaced configuration of roads. We investigated three methods of placing
-seed points:
-\begin{enumerate}
-    \item Random
-    \item Furthest-Point
-    \item Even Streamline Spacing ($\chi^2$ distribution)
-\end{enumerate}
+seed points.
 
+\paragraph{Random}
 Purely random placement of seed points generally results in both clusters of
@@
 
+\paragraph{Furthest-Point}
 In the furthest-point placement, we place each new seed point within the bounds
@@
 candidate seed points which is small compared to the infinite number of
-possible seed points (constrained only by floating point precision) and choose
-the one furthest from all existing streamlines. In practice, this offers a
-significant improvement over the random seed placement, and results in a more
-even distribution of streamlines.
+possible seed points and choose the one furthest from all existing streamlines.
+In practice, this offers a significant improvement over the random seed
+placement, and results in a more even distribution of streamlines.
 
+\paragraph{$\chi^2$ Streamline Spacing}
 However, even choosing well-spaced seed points will not necessarily result in
@@
 space, and perform a chi-squared goodness of fit test on distribution of points
-across bucket, where the expected distribution is uniform:
+across buckets:
 \[
-    \chi^2 = \sum_{i=1}^{n^2} \frac{(O_i-E_i)^2}{E_i}
+    \chi^2 = \sum_{i=1}^{n}\sum_{j=1}^n \frac{(O_{ij}-E_{ij})^2}{E_{ij}}
 \]
 where $n$ is the number of buckets along each axis, $O_i$ is the observed
-number of points in the $i$th bucket, and $E_i=\frac{m}{n^2}$ is the
-expected number of points in each bucket, assuming a uniform distribution,
-where $m$ is the total number of points in the space.
+number of points in the bucket at position $(i,j)$, and $E_i=\frac{m}{n^2}$ is
+the expected number of points in each bucket, assuming a uniform distribution
+and $m$ is the total number of points in the space.
 
@@
 In Section \ref{sec:seedpoints}, we outlined three methods of seedpoint
-placement: random, furthest-point, and $\chi^2$ even spacing. All three
-are implemented by functions in the \texttt{Streamline} module, and follow a
+placement: random, furthest-point, and $\chi^2$ even spacing. All three are
+implemented by functions in the \texttt{Streamline} module, and follow a
 similar template. Extending this module with other methods is as simple as
-implementing a new function with the same type signature as
-\texttt{placeSeeds}, adding a new value constructor to the datatype
+implementing a new function with the same type signature as one of the existing
+seed-placement functions, adding a new value constructor to the datatype
 \texttt{PlacementMethod}, and adding a new case to the function
-\texttt{placeStreamlines} to match the new data constructor. The purity of
+\texttt{placeStreamlines} to match the new value constructor. The purity of
 functions in a language such as Haskell makes it simple to drop in replacements
@@
 stored a set of 2D points and could be queried to determine which of the stored
-points was closest to the query point. This is implemented using a type
-family, with an instance utilizing a blocked 2D array implementation. This
-instance was sufficient for testing purposes, but is currently the performance
-bottleneck in streamline tracing. A new instance of the \texttt{Storage}
-type family could easily be defined in accordance with the type signature to
-provide a more efficient representation. For example, a k-d tree instance of
-the type family could offer higher performance. This well-defined abstraction
-boundary makes this aspect of our pipeline very easily extensible.
+points was closest to the query point. This is implemented using a type family,
+with an instance utilizing a blocked 2D array implementation. This instance was
+sufficient for testing purposes, but is currently the performance bottleneck in
+streamline tracing. A new instance of the \texttt{Storage} type family could
+easily be defined in accordance with the type signature to provide a more
+efficient representation. For example, an instance of the type family
+implementing a k-d tree could offer improved performance. This well-defined
+abstraction boundary makes this aspect of our pipeline very easily extensible.
 
@@
 and the mathematical formulas used for calculating eigenvectors in Section
-\ref{sec:eigenvectors}.  The 75-line function
-\texttt{cal\_eigenvecs\_onevert\_quad} in Chen's implementation. While it is
-documented, we found it difficult to determine the side effects it had. In a
-language without mutable state or side effect, we found it significantly easier
-to fully understand the behavior of functions.  In contrast, even well-written
+\ref{sec:eigenvectors}.  While the 75-line function
+\texttt{cal\_eigenvecs\_onevert\_quad} in Chen's implementation is documented,
+we found it difficult to determine its side effects. We found that a functional
+language without mutable state or side effect made it significantly easier to
+fully understand the behavior of functions.  In contrast, even well-written
 code in \cpp or another language in the imperative paradigm can be difficult to
 understand, because the possibility for side effect requires the reader to have
-an understanding of any part of the system that the function can affect, which
-in practice can make the effort to understand any small part of a large
-codebase much greater.
+an understanding of any part of the system that the function can affect. In
+practice can make the effort to understand any small part of a large codebase
+much greater.
 
@@
 implementation of the work described in \cite{chen}, with the goals being that
-a functional implementation would readily express the mathematics, provide
+a functional implementation would readily express the mathematics, and provide
 modularity that the imperative implementation provide in \cite{chen} could not.
-We believe that, due to the focus on expressibility and modularity, assessed in
-the previous section, that an individual looking to understand the pipeline
-will find our code undestandable and approachable, and lend itself to be
-easily extended. We hope that our work will motivate a greater use of
-functional languages in future efforts in the research of new procedural
-algorithms.
+We believe that, due to the focus on expressibility and modularity that an
+individual looking to understand the pipeline will find our code undestandable
+and approachable, and lend itself to be easily extended. We hope that our work
+will motivate a greater use of functional languages in future efforts in the
+research of new procedural algorithms.
 

### 2013-12-13 JSONParser.hs
some minor style fixes

@@ -1,21 +1,17 @@
-{-# LANGUAGE OverloadedStrings, DeriveGeneric, NamedFieldPuns #-}
+{-# LANGUAGE OverloadedStrings, DeriveGeneric #-}
 
-module JSONParser
-where
+module JSONParser where
 
 import Data.Aeson
 import GHC.Generics
 import Constraint
 import Vector2
 
-data Input = Input
-           {
-            tycon :: String,
-            posx  :: Float,
-            posy  :: Float,
-            dir   :: Float,
-            mag   :: Float
-            } deriving (Show, Generic)
+data Input = Input {  tycon :: String,
+                      posx  :: Float,
+                      posy  :: Float,
+                      dir   :: Float,
+                      mag   :: Float  } deriving (Show, Generic)
 
 instance FromJSON Input
 instance ToJSON Input

### 2013-12-13 Main.hs

@@ -23,15 +23,15 @@ main = getArgs >>= parseCmdArgs
 def :: String
 def = "furthest"
 
-parseCmdArgs ["-h"]            = usage                                   >> exit
-parseCmdArgs ["-v"]            = version                                 >> exit
-parseCmdArgs [file, fw, fh]    = (buildViz file (read fw) (read fh) def) >> exit
-parseCmdArgs [file, fw, fh, o] = (buildViz file (read fw) (read fh) o)   >> exit
-parseCmdArgs _                 = usage                                   >> exit
+parseCmdArgs ["-h"]            = usage                                 >> exit
+parseCmdArgs ["-v"]            = version                               >> exit
+parseCmdArgs [file, fw, fh]    = buildViz file (read fw) (read fh) def >> exit
+parseCmdArgs [file, fw, fh, o] = buildViz file (read fw) (read fh) o   >> exit
+parseCmdArgs _                 = usage                                 >> exit
 
 usage   = putStrLn "usage: tensor <inputFile> <fieldWidth> <fieldHeight> [opt]"
 version = putStrLn "Tensor v0.1"
-exit    = exitWith ExitSuccess
+exit    = exitSuccess
 die     = exitWith (ExitFailure 1)
 
 buildViz :: FilePath -> Float -> Float -> String -> IO ()

### 2013-12-13 NearestNeighbor.hs

@@ -5,15 +5,13 @@ import Data.Vector ((//), (!))
 import qualified Util
 import qualified Vector2 as V2
 
---data Point = Point Float Float deriving Show
-
 class NearestNeighbor a where
   data (Storage a)
   new    :: Float -> Float -> Int -> Storage a
   insert :: Storage a -> (V2.Vector2, a) -> Storage a
   lookup :: Storage a -> V2.Vector2 -> Maybe (V2.Vector2, a)
 
-{-
+{- Naive implementation for testing
 instance NearestNeighbor a where
   data (Storage a) = List [(V2.Vector2, a)]
   new _ _ _ = List []
@@ -30,8 +28,8 @@ instance NearestNeighbor a where
 instance NearestNeighbor a where
   data (Storage a) = Buckets { width   :: Float, height :: Float,
                                xsize   :: Float, ysize  :: Float,
-                               buckets :: V.Vector (V.Vector [(V2.Vector2, a)]) }
-                               deriving Show
+                               buckets :: V.Vector (V.Vector [(V2.Vector2, a)])
+                               } deriving Show
   new    = bucketsNew
   insert = bucketsInsert
   lookup = bucketsLookup

### 2013-12-13 Tensor.hs

@@ -12,7 +12,7 @@ data Tensor = Tensor { a :: Float, b :: Float, d :: Float }
 trivial :: Tensor
 trivial = Tensor 0 0 0
 
--- Creating Tensors from various types of constraints
+-- Creating tensors from various types of constraints
 
 fromRTheta :: Float -> Float -> Tensor
 fromRTheta r t = Tensor (r*cos(2*t)) (r*sin(2*t)) (-1*r*cos(2*t))
@@ -22,7 +22,7 @@ fromXY x y = let xy = -2 * x * y
                  diffSquares = y * y - x * x
              in Tensor diffSquares xy (-1 * diffSquares)
 
--- Basic Tensor operations
+-- Basic tensor operations
 
 add :: Tensor -> Tensor -> Tensor
 add (Tensor a1 b1 d1) (Tensor a2 b2 d2) =
@@ -44,10 +44,8 @@ isReal (Tensor a b c) =
 -- returns a tuple of the (larger, smaller) eigenvectors of a tensor
 eigenvalues :: Tensor -> (Float, Float)
 eigenvalues (Tensor a b d) = 
---  let evals f = (f (a + d) (sqrt (4*b*b + (a - d)*(a - d)))) / 2
   let eval = sqrt (a*a + b*b) 
   in  (eval, (-1)*eval)
--- in  (evals (+), evals (-))
 
 -- returns a tuple of the (major, minor) eigenvectors for a tensor
 eigenvectors :: Tensor -> (V2.Vector2, V2.Vector2)
@@ -58,5 +56,3 @@ eigenvectors (Tensor a b d) =
     let (eval1, eval2) = eigenvalues (Tensor a b d)
         mkEvec eval    = V2.Vector2 1 ((eval - a) / b)
     in (mkEvec eval1, mkEvec eval2)
-
-
